A Phoneme-weighted Levenshtein Distance: A Measure of Cross-linguistic Phonological Similarity

### 2021-02-04 Rmd/te_manuscript.Rmd
Rename Rmd folder

library(dplyr)
library(tidyr)
library(tibble)
library(stringr)
library(ggplot2)
library(here)
library(data.table)
library(tidybayes)
library(gt)
library(lubridate)

participants <- fread(here("Data", "01_participants.csv"), na.strings = "") %>% as_tibble() %>% mutate(date = as_date(date))
dat <- fread(here("Data", "01_processed_coded.csv"), na.strings = "") %>% as_tibble()
posterior <- fread(here("Results", "posterior_draws.csv"), na.strings = "") %>% as_tibble()
predictions <- fread(here("Results", "posterior_predictions.csv"), na.strings = "") %>% as_tibble()
fit <- readRDS(here("Results", "fit3_phon.rds"))
comparison <- readRDS(here("Results", "model_comparison.rds"))
item_effects <- fread(here("Results", "item_effects.csv"), na.strings = "") %>% as_tibble()
group_effects <- fread(here("Results", "group_effects.csv"), na.strings = "") %>% as_tibble()
corrs <- fread(here("Results", "corrs.csv"), na.strings = "") %>% as_tibble()

@@

The experiment was implemented online using Psychopy/Pavlovia [@peirce2019]. Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from a browser (Chrome or Mozilla). First, participants were informed about the aims of the study and gave informed consent for participating. Second, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study. Third, participants completed the experimental task. Before the task, participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants). They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible. English participants were randomly assigned to the list of Catalan or Spanish trials. Spanish participants completed the list of Catalan trials.

### Design

Each trial started with a yellow fixation dot presented during one second on the centre of the screen over a black background. After one second, the audio started playing while the dot remained being displayed until the audio offset. Upon the offset of the fixation point and audio, participants were prompted to write their answer by a ">" symbol. Typed letters were displayed in the screen in real time to provide visual feed-back to participants. Participants were allowed to correct their answer. Then, participants pressed the RETURN key to start and new trial. We excluded trials where participants did not type an existing word in the correspondent language, or did not type anything at all. Typos were allowed, however. Participants contributed a total of `r distinct(dat, participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` valid trials (`r filter(dat, test_language=="Catalan") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Catalan, `r filter(dat, test_language=="Spanish") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Spanish). The task took approximately 15 minutes to be completed.

@@

@@

All analyses were performed in R environment [@rcoreteam2019]. We used the `tidyverse` family of R packages to process data and to generate figures, the `mice` R package [@vanbuuren2013] to impute missing data, and the `brms` R package to [@burkner2017] to fit and compare models, which uses the Stan probabilistic programming language language [@carpenter2016].

>>>>

library(tidyverse)
library(data.table)
library(tidybayes)
library(gt)
library(lubridate)
library(here)

participants <- readRDS(here("Results", "participants.rds"))
stimuli <- readRDS(here("Results", "stimuli.rds"))
accuracy <- readRDS(here("Results", "accuracy.rds"))
fit <- readRDS(here("Results", "fit3_phon.rds"))
comparison <- readRDS(here("Results", "model_comparison.rds"))

@@

The experiment was implemented online using Psychopy/Pavlovia [@peirce2019]. Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from a browser (Chrome or Mozilla). First, participants were informed about the aims of the study and gave informed consent for participating. Second, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study. Third, participants completed the experimental task. Participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants). They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible. English participants were randomly assigned to the list of Catalan or Spanish trials. Participants in the Catalan list were presented with `r count(stimuli, language)$n[1]` trials, and participants in the Spanish list were presented with `r count(stimuli, language)$n[2]` trials.

Each trial started with a yellow fixation dot presented during one second on the centre of the screen over a black background. After one second, the audio started playing while the dot remained being displayed until the audio ended. Upon the offset of the fixation point and audio, participants were prompted to write their answer by a ">" symbol. Typed letters were displayed in the screen in real time to provide visual feed-back to participants. Participants were allowed to correct their answer. Then, participants pressed the RETURN key to start and new trial. We excluded trials where participants did not type an existing word in the correspondent language, or did not type anything at all. Trials where the response was mistyped by only one character were counted as correct, as long as the respond did not correspond to a distinct word. Participants contributed a total of `r distinct(dat, participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` valid trials (`r filter(dat, test_language=="Catalan") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Catalan, `r filter(dat, test_language=="Spanish") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Spanish). The task took approximately 15 minutes to be completed.

@@

Participants listened to one audio file in each trial. This audio file corresponded to a word in Catalan (for Spanish speakers, and for English participants allocated in the Catalan condition) or Spanish (for English speakers allocated in the Spanish condition). The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infancia of Universitat Pompeu Fabra (Barcelona, Spain). These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner. Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@boersma2018]. The average duration of the audios was `r printnum(mean(stimuli$duration), digits = 2)` (*SD* = `r printnum(sd(stimuli$duration), digits = 2)`, *Min* = `r printnum(min(stimuli$duration), digits = 2)`, *Max* = `r printnum(max(stimuli$duration), digits = 2)`). The average duration of the Catalan audios was `r printnum(mean(stimuli[stimuli$language=="Catalan",]$duration), digits = 2)` seconds (*SD* = `r printnum(sd(stimuli[stimuli$language=="Catalan",]$duration), digits = 2)`, *Min* = `r printnum(min(stimuli[stimuli$language=="Catalan",]$duration), digits = 2)`, *Max* = `r printnum(max(stimuli[stimuli$language=="Catalan",]$duration), digits = 2)`), and the average duration of the Spanish audios was `r printnum(mean(stimuli[stimuli$language=="Spanish",]$duration), digits = 2)` seconds (*SD* = `r printnum(sd(stimuli[stimuli$language=="Spanish",]$duration), digits = 2)`, *Min* = `r printnum(min(stimuli[stimuli$language=="Spanish",]$duration), digits = 2)`, *Max* = `r printnum(max(stimuli[stimuli$language=="Spanish",]$duration), digits = 2)`). The average lexical frequency of the words was `r printnum(mean(stimuli$frequency_zipf), digits = 2)` (*SD* = `r printnum(sd(stimuli$frequency_zipf), digits = 2)`, *Min* = `r printnum(min(stimuli$frequency_zipf), digits = 2)`, *Max* = `r printnum(max(stimuli$frequency_zipf), digits = 2)`). The average lexical frequency of the Catalan audios was `r printnum(mean(stimuli[stimuli$language=="Catalan",]$frequency_zipf), digits = 2)` seconds (*SD* = `r printnum(sd(stimuli[stimuli$language=="Catalan",]$frequency_zipf), digits = 2)`, *Min* = `r printnum(min(stimuli[stimuli$language=="Catalan",]$frequency_zipf), digits = 2)`, *Max* = `r printnum(max(stimuli[stimuli$language=="Catalan",]$frequency_zipf), digits = 2)`), and the average lexical frequency of the Spanish audios was `r printnum(mean(stimuli[stimuli$language=="Spanish",]$frequency_zipf), digits = 2)` seconds (*SD* = `r printnum(sd(stimuli[stimuli$language=="Spanish",]$frequency_zipf), digits = 2)`, *Min* = `r printnum(min(stimuli[stimuli$language=="Spanish",]$frequency_zipf), digits = 2)`, *Max* = `r printnum(max(stimuli[stimuli$language=="Spanish",]$frequency_zipf), digits = 2)`). Lexical frequencies in Catalan and Spanish were extracted from SUBTLEX-CAT [@boada2013] and SUBTLEX-ESP [@cuetos2011], respectively, and are expressed as Zipf scores [@van-heuven2012; @zipf1949].

@@

All analyses were performed in R environment [@rcoreteam2019]. We used the `tidyverse` family of R packages to process data and to generate figures, the `mice` R package [@vanbuuren2013] for multiple imputation of missing data, and the `brms` R package [@burkner2017] to fit and compare models, which uses the Stan probabilistic programming language language [@carpenter2016]. Audio duration, intensity and pitch were extracted using the `PraatR` R package [@albin2014].
